---
title: "hw3"
author: "Gage Clawson, Ben Edwards"
date: "5/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


```{r}
# read in data
hw3_raw <- read_csv("HW3Data.csv")
```


1. Create a linear probability model that predicts a respondent’s probability of voting ‘yes’ on the ballot based on their age, income, NEP score, the risk reduction offered by the program, and the cost of the program to that respondent. Show the model and interpret the regression coefficients.

```{r}
p_yes <- lm(vote ~ age + income + NEP + risk + bid, data = hw3_raw)
summary(p_yes)
```

$P_{yes} = 0.1197 + (0.0204*agetothirty) + (-0.0201*agetoforty) + (0.0099*agetofifty) + (-0.0162*agetosixty) + (0.0027*poor) + (0.0075*rich) + (0.0468*veryrich) + (0.0088*onepercent) + (0.0158*NEP) + (0.0007*risk) + (-0.0011*bid)$

- Compared to those that are over the age of sixty (our reference level for the age variable), those from age 18 - 30 are 2% more likely to vote yes, those that are from age 31 - 40 are 2% less likely to vote yes, those that are from age 41 - 50 are 0.99% more likely to vote yes, those that are from age 51 - 60 are 1.6% less likely to vote yes. Compared to the middle class, those that are poor are 2% more likely to vote yes, those that are rich are .7% more likely to vote yes, those that are very rich are 4.7% more likely to vote yes, and those that are in the one percent are 0.88% more likely to vote yes. For every unit of NEP, probability of voting yes increases by 1.6%, for every unit increase of risk, probability of voting yes increases by 0.07%, and for every dollar increase in cost (bid), probability of a yes vote decreases by 0.1%.

2. Based on this regression, what can you say about the value of a single prevented whale death? (Hint: Think about how risk reduction for all whales translates into the number of whale deaths avoided)

```{r}
#save coefficients 
agetothirty <- p_yes$coefficients[5]
agetoforty <- p_yes$coefficients[3]
agetofifty <- p_yes$coefficients[2]
agetosixty <- p_yes$coefficients[4]
intercept <- p_yes$coefficients[1]
one_percent <- p_yes$coefficients[6]
very_rich <- p_yes$coefficients[9]
rich <- p_yes$coefficients[8]
poor <- p_yes$coefficients[7]
NEP <- p_yes$coefficients[10]
risk <- p_yes$coefficients[11]
bid <- p_yes$coefficients[12]

#20/5 = x%/1
#4 = x/1
single_whale_risk_red <- 4

#use first person in dataset to do calculation 
singlewhale <- (risk*hw3_raw$risk[1] - (risk*hw3_raw$risk[1] + (single_whale_risk_red*risk)))*(1/bid)
#The value of a single prevent whale death is $2.7834. 

```


3. Pick three arbitrary respondents. For each of those respondents, estimate their WTP for a VSR program offering 60% risk reduction. 

$(0.001*bid) = 0.1197 + (-0.714*voteyes) + (0.0204*agetothirty) + (-0.0201*agetoforty) + (0.0099*agetofifty) + (-0.0162*agetosixty) + (0.0027*poor) + (0.0075*rich) + (0.0468*veryrich) + (0.0088*onepercent) + (0.0158*NEP) + (0.0007*risk)$

```{r}
mean_pyes <- predict(p_yes, newdata = hw3_raw, type = "response")
mean_pyes <- sum(mean_pyes)/500 #mean probability of yes is .714 

#save coefficients by solving for bid 
#agetothirty_bid <- agetothirty/-bid
#agetoforty_bid <- agetoforty/-bid
#agetofifty_bid <- agetofifty/-bid
#agetosixty_bid <- agetosixty/-bid
#poor_bid <- poor/-bid
#rich_bid <- rich/-bid
#veryrich_bid <- very_rich/-bid
#onepercent_bid <- one_percent/-bid
#nep_bid <- NEP/-bid
#risk_bid <- risk/-bid
#intercept_bid <- intercept/-bid

#select 3 random rows from dataset
#random3 <- sample_n(hw3_raw, 3)
#random3

# A tibble: 3 x 6
#   risk   bid  vote   NEP income      age     
#  <int> <int> <int> <int> <chr>       <chr>   
#1    20    60     1    41 one_percent tosixty 
#2    60    60     1    53 one_percent tofifty 
#3    20    60     1    43 poor        tothirty

#use mean prob of yes for vote variable

#respondant1: risk = 20, bid =?, vote = 0.714, NEP = 41, one_percent, tosixty
#bid_value_1 = (intercept_bid - mean_pyes + agetosixty_bid + onepercent_bid + 41*nep_bid + risk_bid*20)
bid_value_1 <- (intercept - mean_pyes + agetosixty + one_percent + 41*NEP + risk*20)/-bid
#59.5

#respondant2: risk = 60, bid =?, vote = 0.714, NEP = 53, one_percent, tofifty
bid_value_2 <- (intercept - mean_pyes + agetofifty + one_percent + 53*NEP + risk*60)/-bid
#290

#respondant3: risk = 20, bid =?, vote = 0.714, NEP = 43, poor, tothirty
bid_value_3 <- (intercept - mean_pyes + agetothirty + poor + 43*NEP + risk*20)/-bid
#118

#will put this table into the appendix: 
#   risk   WTP    vote  NEP   income       age     
#  <int> <int>   <int> <int>   <chr>       <chr>   
#1    20    59.5   1    41    one_percent tosixty 
#2    60    290    1    53    one_percent tofifty 
#3    20    118    1    43    poor        tothirty

```


4. Now repeat this process and estimate the mean WTP for a VSR program offering 60% risk reduction among Santa Barbara County households.

```{r}
#find mode of age variable
age_mode <- hw3_raw %>%
  group_by(age) %>%
  tally()
#tothirty is the mode

#find the mode of income 
income_mode <- hw3_raw %>%
  group_by(income) %>%
  tally()
#income mode is one_percent? 

#find nep mean 
nep_mean <- mean(hw3_raw$NEP)
#38.366

#Assume a 60% risk reduction and plug above values into bid eq

bid_mean_sb <- (intercept - mean_pyes + agetothirty + one_percent + nep_mean*NEP + risk*60)/-bid

#82.5
```

The mean WTP for a VSR program offering a 60% risk reduction among SB residents is $82.5. 

5. If there are 150,000 households in Santa Barbara County, estimate the total benefit of a VSR program to these households.

```{r}
all_sb <- 150000*bid_mean_sb  
#12,375,410
```

The total benefit of a VSR program to 150,000 households in SB is $12,375,410. 

6. Based only on these benefits, do the benefits of the VSR program outweigh the costs?

```{r}
#total costs are 7 mil

all_sb - 7000000
#5375410

```
Net benefits = $12,375,410 - $7,000,000 = $5,375,410. Yes benefits outweigh the costs. 


When a container ship reduces its speed, it not only reduces the risk of colliding with a whale but also emits less carbon per distance traveled. For this reason, the WhaleStrikes Group Project recommended enrolling container ships in voluntary carbon trading markets to mitigate fatal whale strikes. Assume that for any ship transiting the Santa Barbara Channel, a speed reduction (that results in a 60% risk reduction) will cost the shipper $1,000, but will result in 20 fewer tons of CO2 emitted per transit.

7. Suppose the price of a carbon credit is $Z per ton, at what price Z* will the shipping industry decide to adopt the VSR for purely self-interested reasons? (Note: if ships voluntarily reduce speed, then the 7 million cost (noted above) of implementing the VSR would be avoided)

```{r}
#20*Z = 1000
z <- 1000/20

# $50 per ton
```
The shipping industry will decide to adopt the VSR at $50 per ton of CO2. 


8. Now suppose the carbon credit price is exactly Z*, so all ships voluntarily reduce speed to achieve the 60% risk reduction. Approximately how many whales would be saved by this? What would be the social value of allowing ships to enter the carbon trading market?

```{r}
#4/1 = 60/x
# x = 15


nsv <- ((-1000 + 1000) + all_sb) - (singlewhale*150000*15)
#$6,112,760.69417361 
```

a. A risk reduction of 4% saves one whale, which means a risk reduction of 60% will save 15 whales. 
b. 


 


